### 2022-04-10 model specification.Rmd
update model specification binary graph learning

---
title: "Comments on binomial graph learning"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Binary graph learning model 
### Version 1 (Eigenvector matrix from Graph Laplacian)
Let $Y_{i,j}$ denote the measurement on the node $i$ at round $j$, where $j = 1, \dots, M$, and $i = 1, \dots, N$. $Y_{i,j}$ is a binomial signal that can be 1, or 0. Suppose the signals at round $j$ denoted by $Y[, j]$ for all $N$ nodes are independent of the signals at round $k$ denoted by $Y[,k ]$, for $i \neq k$. Let $p_{i,j}$ denote the probability of $Y_{i,j} = 1$. Our model assumes
$$\text{logit}(p_{i,j}) = \alpha_j + (\chi h )_i,$$

where $\chi$ is the eigenvector matrix from Graph Laplacian $L$, $h$ is a vector of latent factors that governs $p_{i, j}$ through $\chi$, and $\alpha_j$ is a round specific parameter at round $j$.

### Version 2 (Adjacency matrix from Graph)
#### Graph 
We consider a weighted undirected graph $G = (V, E)$, with the vertices set $V = {1, 2, \dots, N}$, and edge set $E$.  Let $\mathbf{A}$ denote the weighted adjacency matrix for the graph $G$. In the case of weighted undirected graph, $\mathbf{A}$ is a square and symmetric matrix.

#### Signals on the graph
Let $Y_{i,j}$ denote the signal on the node $i$ of graph $G$ at round $j$, where $j = 1, \dots, M$, and $i = 1, \dots, N$. We assume that $Y_{i,j}$ is a binary signal that can be 1, or 0. Suppose the signals at round $j$ denoted by $Y[, j]$ for all $N$ nodes are independent of the signals at round (or strata) $k$ denoted by $Y[,k ]$, for $i \neq k$, borrowing the idea of conditional logistic regression. Let $p_{i,j}$ denote the probability of $Y_{i,j} = 1$. Our model assumes

\begin{equation}
\label{eq:binaryglm}
\text{logit}(p_{i,j}) = \alpha_j + (\mathbf{A} h )_i,
\end{equation}

where $\mathbf{A}$ is the adjacency matrix from the graph $G$, $h$ is a vector of latent factors that governs $p_{i, j}$ through $\mathbf{A}$, and $\alpha_j$ is a round specific parameter at round or strata $j$. In the following, we will use

## Method of Estimation

The conditional likelihood function for model (\ref{eq:binaryglm}) given  is (I will update soon)

<!-- \begin{equation} -->
<!-- 	\label{eq:dongmodel} -->
<!-- 	\begin{aligned} -->
<!-- 		\min_{L \in \mathbb{R}^{N\times N}, \phantom{..} Y \in \mathbb{R}^{N\times P}} ||X&-Y||_F^2 + \alpha \text{tr}(Y^TLY) + \beta||L||_F^2 \\ -->
<!-- 		\textrm{s.t.}& \quad \text{tr}(L) = N, \\ -->
<!-- 		&\quad L_{i,j} = L_{j,i} \leq 0, \phantom{..} i \neq j, \\ -->
<!-- 		&\quad L\cdot \textbf{1} = \textbf{0} \\ -->
<!-- 	\end{aligned} -->
<!-- \end{equation} -->

## Miscellaneous 
Constraints:

Sigmoid computation,
Bound and Branch method of optimization

hint for factor analysis solution might be useful because there is another object function varimax.

https://conservancy.umn.edu/bitstream/handle/11299/95957/Choi_umn_0130E_11451.pdf?sequence=1&isAllowed=y

https://web.stanford.edu/~boyd/papers/pdf/max_sum_sigmoids.pdf

* complete separation do not exist. observable metric ?, it is a potential new topic to be able to identify whether complete separation exists. possible discuss it in the Discussion and future work.

* While working on this one, you may also consider implementing our regressor paper.

## Model Specification

## Our goal is to estimate the Graph Laplacian

* We do not have to use the eigenvector matrix $\chi$. Instead, we may consider adjacency matrix $A$ in place of $\chi$. This is a viable direction.

Maximum Likelihood estimation, Quasi likelihood estimation,