### 2022-04-27 poster html copy/poster html.Rmd
Poster update.

---
title: Graph Topology Learning from Binary Signals on Graph
author:
  - name: Skip Moses
affiliation:
    address: Department of Mathematics and Statistics, California State University, Chico
column_numbers: 4
primary_colour: "#9D2235"
secondary_colour: "#FFFFFF"
poster_height: "36in"
poster_width: "60in"
# logoright_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
# logoleft_name: https&#58;//raw.githubusercontent.com/brentthorne/posterdown/master/images/betterhexlogo.png
output: 
  posterdown::posterdown_html:
    self_contained: false
bibliography: ref.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

Data often has  an underlying structure or geometry that can be modeled as a signal on the vertices of a weighted, undirected graph. Traditionally, the emphasis was on using a underlying graph, or network, to understand the properties of signals over the vertices. Recently, there has been a surge in converse problem; of learning a graph structure from a set of signals satisfying some constraints [@xia2021graph] [@dong2016learning] [ @ortega2018graph]. In previous research [@dong2016learning], [@tugnait2021sparse], [@pu2021learning], and [@saboksayr2021accelerated], signals on graph were assumed to follow multivariate Gaussian distributions, but there has been little exploration in learning a network from binomial signals. In this work, we develop a novel methodology that allows for learning graph topology given a set of binary signals on the graph. An example of such signals on graph is presented in Figure 1.

```{r GTG, echo=FALSE,out.width="90%", out.height="20%",fig.cap="Circular Embedding of ground truth graph",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("pictures/gt_graph.png"))
```

# Methods

#### Signals on the graph and model specification
We consider a weighted undirected graph $G = (V, E)$, with the vertices set $V = {1, 2, \dots, N}$, and edge set $E$.  Let $\mathbf{A}$ denote the weighted adjacency matrix for the graph $G$. In the case of weighted undirected graph, $\mathbf{A}$ is a square and symmetric matrix.

Let $Y_{i,j}$ denote the signal on the node $i$ of graph $G$ at round $j$, where $j = 1, \dots, M$, and $i = 1, \dots, N$. We assume that $Y_{i,j}$ is a binary signal that can be 1, or 0. 
Suppose the signals at stratum $j$ denoted by $Y[, j]$ for all $N$ nodes are independent of the signals at stratum $k$ denoted by $Y[,k ]$, for $j \neq k$. Let $p_{i,j}$ denote the probability of $Y_{i,j} = 1$. Our model assumes

\begin{equation}
\label{eq: binaryglm}
\text{logit}(p_{i,j}) = (\mathbf{A} h )_i,
\end{equation}

where $\mathbf{A}$ is the adjacency matrix from the graph $G$, $h$ is a vector of latent factors that governs $p_{i, j}$ through $\mathbf{A}$ and assumed to be a standard normal random vector.

#### Method of Estimation
##### Maximum likelihood for one stratum

Consider the probability mass function for a given $A$ and signal $y = Y[,k]$
\begin{align}
P_{Ah}(y_i) &= p^{y_i}(1-p)^{1-y_i} \\
            &= \left(\frac{e^{A[i,]h}}{1 + e^{A[i,]h}}\right)^{y_i}\left(1- \frac{e^{A[i,]h}}{1 + e^{A[i,]h}}\right)^{1-y_i} \\
            &= \left(\frac{e^{A[i,]h}}{1 + e^{A[i,]h}}\right)^{y_i}\left(\frac{1}{1 + e^{A[i,]h}}\right)^{1-y_i}   \\
            &= \frac{e^{y_iA[i,]h}}{1 + e^{A[i,]h}}

\end{align}

Therefore, our Likelihood function will be given by

\begin{align}
\mathcal{L}(h) = \prod_{i=1}^N\left(\frac{e^{y_iA[i,]h}}{1 + e^{A[i,]h}}\right)
\end{align}

In order to maximize we consider the natural logarithm of our likelihood

\begin{align}
\log(\mathcal{L}(h)) = \sum_{i=1}^N\left(y_i(A[i,]h) - \log(1 + e^{A[i,]h})\right)
\end{align}

##### Optimization Program

Taking inspiration from the above derivation we will solve for estimated $A$ by maximizing the following

\begin{equation}
\label{eq:optimization}
\begin{aligned}
&\max_{A,h} \sum_{j = 1}^M\sum_{i=1}^N\left(y_{i,j}(A[i,]h) - \log(1 + e^{A[i,]h})\right) - \alpha \vert L \vert _ F \\
\textrm{s.t.} \quad  &A_{i,j} = 0 \text{ if } i =j\\
                     &A_{i,j} \geq 0 \text{ if } i \neq j\\
                     &\mathbb{1}h^T = 0 \\
                     &\max(h) \leq a \\
                     &\min(h) \geq b
\end{aligned}
\end{equation}
where $\alpha$ is a tuning parameter for controlling the sparsity and $a$ and $b$ are tuning parameters for restricting the spread of the values of $h$.

<hr /> 
<hr /> 
$\textbf{Algorithm 1:}$ Estimate $A$ given $Y$.
<hr /> 
<hr />

\begin{align*}
&1) \textbf{ Input: } \text{ Input a signal } Y.\\
&2) \textbf{ Output: } \text{ Output an estimated } A.\\
&3) \textbf{ Initialization: } h_{i,0} \sim \mathcal{N}(0,1) \text{ for } i = 1,\ldots, N \\
&4) \textbf{ for } t = 1, \ldots, iter:\\
&5) \textbf{ Update } A \textbf{ given } h:\\
&6) \textbf{        Update } h \textbf{ given } A:\\
&7) \textbf{ end for }
\end{align*}
<hr />

# Results

In each iteration of step 5) and 6) of Algorithm 1, the optimization program gives a Disciplined Concave Program that can be solved efficiently in Python with the CVXPY library. We use a random Erd&#337;s-Reyni graph on $N = 20$ nodes and edge probability $p = 0.2$. Each edge weight $w_i \sim \mathcal{U}(0,1)$.

##### Synthetic Data Generation

To generate a single signal for a given adjacency matrix $A \in \mathbb{R}^{N\times N}$ and latent factor vector $h \in \mathbb{R}^N$ (each $h_i \sim \mathcal{N}(0,1)$), a vector of odds, $Ah$, is compared to a vector $T$, where $T_i$ selected uniformly at random for each $i = 1, \ldots , N$. If $T_k < Ah_k$, then node $k$ takes the signal value of $1$, and $0$ otherwise. We generate $M = 100$ synthetic signals, and use them to learn an adjacency matrix. The learned graph is found with parameter values $\alpha = .2$, $a = 1$ and $b = -1$.

Heat maps of the ground truth adjacency matrix and estimated adjacency matrix are given in Figure 2, along with circular embedding of the ground truth and estimated graphs in Figure 3.

```{r HM, echo=FALSE,out.width="49%", out.height="20%",fig.cap="(Left) A heatmap of the ground truth adjacency matrix. (Right) A heatmap of the estimated adjacency matrix.",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("pictures/gt_A_heatmap.png","pictures/est_A_heatmap.png"))
```

```{r EAG, echo=FALSE,out.width="90%", out.height="20%",fig.cap="Circular Embedding of Estimated graph",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("pictures/est_graph.png"))
```

# Discussion

Currently, GSP frameworks such as ours have limitations. For instance, these frameworks require the input to be set of complete signals. This limits the size of graphs that can be learned. Another related limitation, is the scalability of the algorithms used to learn a topology for the graph. In @saboksayr2021accelerated the author shows fast proximal-gradient iterations can be applied to the framework given by @kalofolias2016learn converges to a globally optimal solution in $O (1/k)$. A natural next step would be to apply fast proximal-gradient iterations our framework to overcome scalability issues.

```{r, include=FALSE}
knitr::write_bib(c('knitr','rmarkdown','posterdown','pagedown'), 'packages.bib')
```
# Acknowledgement
This work is supported by Research, Scholarship, and Creative Activity (RSCA) Award 2021-2022 through Chico State Enterprises. This work would not have been possible without the guidence of Dr. Jing Guo. I am extremely grateful for all of meetings over the semester.

# References