### 2022-04-28 poster html copy/poster html.Rmd
Updated logos.

---
title: Graph Topology Learning from Binary Signals on Graph
author:
  - name: Skip Moses
  - name: Jing Guo
affiliation:
    address: Department of Mathematics and Statistics, California State University, Chico
column_numbers: 4
primary_colour: "#9D2235"
secondary_colour: "#FFFFFF"
poster_height: "36in"
poster_width: "60in"
logoright_name: pictures/DSI_Hexbin_sticker_01.png
#logoright_name: pictures/logo1.png
output: 
  posterdown::posterdown_html:
    self_contained: false
bibliography: ref.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

Data often has  an underlying structure or geometry that can be modeled as a signal on the vertices of a weighted, undirected graph. Traditionally, the emphasis was on using a underlying graph, or network, to understand the properties of signals over the vertices. Recently, there has been a surge in converse problem; learning a graph structure from a set of signals satisfying some constraints [@dong2016learning] [ @ortega2018graph]. In previous research [@dong2016learning], [@tugnait2021sparse], and [@saboksayr2021accelerated], signals on graph were assumed to follow multivariate Gaussian distributions, but there has been little exploration in learning a network from binomial signals. In this work, we develop a novel methodology that allows for learning graph topology given a set of binary signals on the graph as shown in Figure \@ref(fig:GTG).

```{r GTG, echo=FALSE,out.width="80%", out.height="25%",fig.cap="Circular Embedding of random Erd&#337;s-Reyni Graph with edge probability p = 0.2. Each edge weight follows uniform random distribution on (0,1). Grey nodes indicate a signal value of 0 and black nodes indicate a signal value of 1.",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("pictures/gt_graph.png"))
```

# Methods

#### Signals on the graph and model specification
We consider a weighted undirected graph $G = (V, E)$, with the vertices set $V = {1, 2, \dots, N}$, and edge set $E$.  Let $\mathbf{A}$ denote the weighted adjacency matrix of $G$. In the case of weighted undirected graph, $\mathbf{A}$ is a square and symmetric matrix.

Let $\mathbf{Y}_{i,j}$ denote the signal on the node $i$ of graph $G$ at round $j$, where $j = 1, \dots, M$, and $i = 1, \dots, N$. We assume that $\mathbf{Y}_{i,j}$ is a binary signal that can be 1, or 0. 
Suppose the signals at stratum $j$ denoted by $\mathbf{Y}[, j]$ for all $N$ nodes are independent of the signals at stratum $k$ denoted by $\mathbf{Y}[,k ]$, for $j \neq k$. Let $p_{i,j}$ denote the probability of $\mathbf{Y}_{i,j} = 1$. Our model assumes

\begin{equation}
\label{eq: binaryglm}
\text{logit}(p_{i,j}) = (\mathbf{A} h )_i,
\end{equation}

where $\mathbf{A}$ is the adjacency matrix from the graph $G$, $h$ is a vector of latent factors that governs $p_{i, j}$ through $\mathbf{A}$ and assumed to be a standard normal random vector.

##### Maximum likelihood for one stratum

Consider the probability mass function for a given $\mathbf{A}$ and signal $y = \mathbf{Y}[,k]$
\begin{align}
P_{\mathbf{A}h}(y_i) &= p^{y_i}(1-p)^{1-y_i} \\
            &= \left(\frac{e^{\mathbf{A}[i,]h}}{1 + e^{\mathbf{A}[i,]h}}\right)^{y_i}\left(1- \frac{e^{\mathbf{A}[i,]h}}{1 + e^{\mathbf{A}[i,]h}}\right)^{1-y_i} \\
            &= \left(\frac{e^{\mathbf{A}[i,]h}}{1 + e^{\mathbf{A}[i,]h}}\right)^{y_i}\left(\frac{1}{1 + e^{\mathbf{A}[i,]h}}\right)^{1-y_i}   \\
            &= \frac{e^{y_i\mathbf{A}[i,]h}}{1 + e^{\mathbf{A}[i,]h}}

\end{align}

Therefore, our Likelihood function will be given by

\begin{align}
\mathcal{L}(h) = \prod_{i=1}^N\left(\frac{e^{y_i\mathbf{A}[i,]h}}{1 + e^{\mathbf{A}[i,]h}}\right)
\end{align}

In order to maximize we consider the natural logarithm of our likelihood

\begin{align}
\log(\mathcal{L}(h)) = \sum_{i=1}^N\left(y_i(\mathbf{A}[i,]h) - \log(1 + e^{\mathbf{A}[i,]h})\right)
\end{align}

##### Optimization Program

Taking inspiration from the above derivation we will solve for estimated $\mathbf{A}$ by maximizing the following

\begin{equation}
\label{eq:optimization}
\begin{aligned}
&\max_{\mathbf{A},h} \sum_{j = 1}^M\sum_{i=1}^N\left(y_{i,j}(\mathbf{A}[i,]h) - \log(1 + e^{\mathbf{A}[i,]h})\right) - \alpha \vert L \vert _ F \\
\textrm{s.t.} \quad  &\mathbf{A}_{i,j} = 0 \text{ if } i =j;\,\, \mathbf{A}_{i,j} \geq 0 \text{ if } i \neq j\\
                     &\max(h) \leq a ;  \,\, \min(h) \geq b \,\, ; \mathbb{1}h^T = 0  
\end{aligned}
\end{equation}
where $\alpha$ is a tuning parameter for controlling the sparsity and $\mathbf{A}$ and $b$ are tuning parameters for restricting the spread of the values of $h$.

We presented our optimization algorithm below. In each iteration of step 5) and 6) of Algorithm 1, the optimization program gives a Disciplined Concave Program that can be solved efficiently in Python with the CVXPY library.

<hr />
\begin{align*}
&\textbf{Algorithm 1: Binary Signal Graph Learning} \\
&1) \textbf{ Input: } \text{ Input a signal } \mathbf{Y}.\\
&2) \textbf{ Output: } \text{ Output an estimated } \mathbf{A}.\\
&3) \textbf{ Initialization: } h_{i,0} \sim \mathcal{N}(0,1) \text{ for } i = 1,\ldots, N \\
&4) \textbf{ for } t = 1, \ldots, iter:\\
&5) \textbf{ Update } \mathbf{A} \textbf{ given } h:\\
&6) \,\,\,\, \textit{ Fix }  h \textit{ in Optimization Program and solve } \mathbf{A} \\
&7) \textbf{        Update } h \textbf{ given } \mathbf{A}:\\
&8) \,\,\,\, \textit{ Fix } \mathbf{A} \text{ in Optimization Program and solve h} \\
&9) \textbf{ end for }
\end{align*}
<hr />

# Results
##### Synthetic Data Generation
Let $\mathbf{A}$ be the adjacency matrix of the graph in Figure \@ref(fig:GTG), and $h \in \mathbb{R}$ be such that $h_i \sim \mathcal{N}(0,1)$. Set $p$ as a logistic function of $\mathbf{A}h$, and compare $p$ to a vector $t$, where each $t_i$ follows a random uniform distribution $\mathcal{U}(0,1)$. If $t_k < p_k$, then node $k$ takes the signal value of $1$, and $0$ otherwise. We generate $M = 100$ synthetic signals, and use them to learn an adjacency matrix.

##### Graph Learning Results

The learned graph is found with parameter values $\alpha = .2$, $a = 1$ and $b = -1$ by implementing Algorithm 1. We can see the learned adjacency matrix shares some similar features to the ground truth (Figure \@ref(fig:HM)), but the learned adjacency matrix is sparser than the ground truth (Figure \@ref(fig:HM) and Figure \@ref(fig:EAG)).

```{r HM, echo=FALSE,out.width="49%", out.height="20%",fig.cap="(Left) A heatmap of the ground truth adjacency matrix. (Right) A heatmap of the estimated adjacency matrix. ",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("pictures/gt_A_heatmap.png","pictures/est_A_heatmap.png"))
```

```{r EAG, echo=FALSE,out.width="80%", out.height="25%",fig.cap="Circular Embedding of Estimated graph. Note the overall weight of each edge is smaller than the edge weights of the ground truth.",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("pictures/est_graph.png"))
``` 
# Discussion
To our knowledge, this is the first work on learning graph topology from binary signals. We have developed a method to effectively learn the graph adjacency matrix from binary signals. This model can be easily extended for learning binomial signals in general. Incorporating parameter optimization techniques would be a natural improvement on this model. Moreover, it would be desirable to develop scalable algorithms that can be used to learn a topology for a large graph from signals. Recently, in @saboksayr2021accelerated the author shows fast proximal-gradient iterations can be applied to the framework given by @kalofolias2016learn improving the over-all runtime.

```{r, include=FALSE}
knitr::write_bib(c('knitr','rmarkdown','posterdown','pagedown'), 'packages.bib')
```
# Acknowledgement

 <body>      
        <p style="color:black;font-size:24px;">This work is supported by Research, Scholarship, and Creative Activity (RSCA) Award 2021-2022 through Chico State Enterprises. Thank you to Robin Donatello for approving this work for Data Science Capstone Project.</p>    
</body>

```{r , echo=FALSE,out.width="60%", out.height="10%",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("pictures/logo1.png"))
```

# References